---
title: "PCaMM trials - HPG-Hg(II) interactions"
author: "Benjamin D. Peterson"
date: "2024-06-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(readxl)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(echo = TRUE)
new.wd <- gsub("\\/[[:alpha:]]+$", "", getwd())
knitr::opts_knit$set(root.dir = new.wd)
rm(new.wd)
```

# Read in metadata

First I prepped the metadata for this specific experiment and selected what I needed.

```{r prep_metadata, message = FALSE}
incubation_metadata <- read_xlsx("metadata/PCaMM_incubations_metadata.xlsx",
                                 sheet = "incubationInfo",
                                 skip = 1) %>%
  filter(incubationType == "PCaMM trial runs - HPG-Hg(II) interactions") %>%
  mutate(start_point = lubridate::ymd_hm(paste(startDate, startTime))) %>%
  select(incubationID, treatment, start_point)

# Sediments
sediment_container_info <- read_xlsx("metadata/PCaMM_incubations_metadata.xlsx",
                                 sheet = "containerInfo_Hg",
                                 skip = 1) %>%
  mutate(end_point = lubridate::ymd_hm(paste(dateKilled, timeKilled)))

sediment_metadata <- right_join(sediment_container_info,
                                incubation_metadata) %>%
  select(containerID, incubationID, depth_cm, treatment, start_point, end_point)

# Porewaters
PW_container_info <- read_xlsx("metadata/PCaMM_incubations_metadata.xlsx",
                                 sheet = "containerInfo_other") %>%
  mutate(end_point = lubridate::ymd_hm(paste(dateKilled, timeKilled)))
PW_metadata <- right_join(PW_container_info,
                          incubation_metadata) %>%
  select(containerID, incubationID, treatment, Notes, start_point, end_point)

```


# Prepare sediment data

I read in sediment MeHg data. NOTE, THIS IS PRELIMINARY DATA AND HAS NOT BEEN REVIEWED YET.

```{r prep_sed_MeHg_data}
sed_MeHg_data <- read_xlsx("../data/20240531a_MerxM_iCAP_raw_seds_CorrectOrder.xlsx",
                           sheet = "Sample_data") %>%
  mutate(MeHg_amb_ng_g = `Ambient MeHg`,
         MeHg_204_ng_g = `MeHg from Source_1`) %>%
  filter(grepl("PCaMM", containerID)) %>%
  select(containerID, MeHg_amb_ng_g, MeHg_204_ng_g)
```

Prepare the freeze-dry information.

```{r prepare_freeze_dry_info}
percent_dry_weight_data <- sediment_container_info %>%
  mutate(percent_dry_weight = (massWithDriedSample - massNonSampleMaterial) / (massWithWetSample - massNonSampleMaterial) * 100) %>%
  select(containerID, percent_dry_weight)

```

Then I combined all the sediment data.

```{r combine_all_sed_data, message = FALSE}
all_sed_data <- left_join(sediment_metadata,
                          percent_dry_weight_data) %>%
  left_join(sed_MeHg_data %>%
              mutate(containerID = gsub("_0", "_00", containerID)))
```



# Check percent dry weight

```{r plot_sed_dry_weight_data, echo = FALSE}
all_sed_data %>%
  filter(!is.na(percent_dry_weight)) %>%
  ggplot(aes(x = treatment,
             y = percent_dry_weight,
             col = depth_cm,
             shape = incubationID)) +
  geom_point() +
  ylim(c(40, 70)) +
  theme_bw()
```


# Check MeHg concentrations

Noting here again that THIS IS PRELIMINARY DATA THAT HAS NOT BEEN REVIEWED YET.
```{r plot_sed_MeHg_data, echo = FALSE}
ggpubr::ggarrange(all_sed_data %>%
                    filter(!is.na(MeHg_204_ng_g)) %>%
                    ggplot(aes(x = treatment,
                               y = MeHg_amb_ng_g,
                               col = depth_cm,
                               shape = incubationID)) +
                    geom_point() +
                    ylim(c(0, max(all_sed_data$MeHg_amb_ng_g, na.rm = TRUE)*1.1)) +
                    theme_bw() +
                    theme(legend.position = "inside",
                          legend.position.inside = c(0.8, 0.2)),
                  all_sed_data %>%
                    filter(!is.na(MeHg_204_ng_g)) %>%
                    ggplot(aes(x = treatment,
                               y = MeHg_204_ng_g,
                               col = depth_cm,
                               shape = incubationID)) +
                    geom_point() +
                    ylim(c(0, max(all_sed_data$MeHg_204_ng_g, na.rm = TRUE)*1.1)) +
                    theme_bw() +
                    theme(legend.position = "none")) 
```


# Prepare porewater data

I read in porewater MeHg data. NOTE, THIS IS PRELIMINARY DATA AND HAS NOT BEEN REVIEWED YET.

```{r prep_PW_MeHg_data}
PW_MeHg_data <- read_xlsx("../data/20240530a_MerxM_iCAP_raw_waters_204_PCaMM.xlsx",
                          sheet = "Sample_data") %>%
  mutate(MeHg_amb_ng_l = `Ambient MeHg`,
         MeHg_204_ng_l = `MeHg from Source_1`) %>%
  filter(grepl("PCaMM", containerID)) %>%
  select(containerID, MeHg_amb_ng_l, MeHg_204_ng_l)
```

Then I combined all the porewater data.

```{r combine_all_PW_data, message = FALSE}
all_PW_data <- left_join(PW_metadata,
                         PW_MeHg_data)
```

# Visualize PW data

Noting here again that THIS IS PRELIMINARY DATA THAT HAS NOT BEEN REVIEWED YET.
```{r plot_PW_MeHg_data, echo = FALSE}
all_PW_data %>%
  # filter(!is.na(MeHg_204_ng_g)) %>%
  ggplot(aes(x = treatment,
             y = MeHg_204_ng_l,
             col = Notes,
             shape = incubationID)) +
  geom_point() +
  ylim(c(0, max(all_PW_data$MeHg_204_ng_l, na.rm = TRUE)*1.1)) +
  theme_bw()
```